---
title: "Filter startups by country"
author: "Michael Lagunov"
date: "9/5/2020"
output: 
  ioslides_presentation:
    widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Background

This project is a homework of Data Science Specialization by John Hopkins University, course of developing data products. The main goal is to practice new knowledge of making shiny applications.

Current presentation is a complement to the shiny website, which you can find here:

[https://michael-lagunov.shinyapps.io/Filter-startups-by-country](https://michael-lagunov.shinyapps.io/Filter-startups-by-country)

## About project

Running the application, you will be able to see startups separated by country. You have an opportunity to change a country to any you like

Also, it is very interesting to see if there are any specific startup-cluster cities.
For example in Russia, Moscow has almost a half of all the startups, that are in the database. In UK London has nearly 70%

There may be some limitations due to the sample size

## Data workflow {.smaller}

1. Parse startups from open access websites, with `Rvest` library. I used [this website](https://innmind.com). Data were stored in a DB, but converted to CSV for this project. See the last slide for more info

2. Made a shiny application, using ui.R and server.R files. A chunk of server.R file lies below, see the last slide for more info

```{r, echo = TRUE, eval = FALSE}
shinyServer(function(input, output, session) {
    output$companies <- function() {
        comp <-  companies %>% 
            select(name, country, city, address) %>%
            filter(country == input$Country) %>%
            knitr::kable("html") %>%
            kable_styling("striped", full_width = F)
    }
    tbl <- reactive({
        companies %>%
            filter(country == input$Country) %>%
            group_by(city) %>%
            summarize(freq = n(), .groups = "drop") %>%
            arrange(desc(freq))
    })
    output$plot <- renderPlotly({
        plot_ly(data = tbl(), values = ~freq, labels = ~city, type = "pie")
    })
})
```

## Resources

* [Github with shiny and presentation code](https://github.com/lagunovmike/datasciencecoursera/tree/master/developing-data-products/Assignment-w4)
* [Parsing script](https://github.com/lagunovmike/Sandbox/blob/master/Startups-info/Parse_companies.R)
* Contact me at lagunovmike@yahoo.com